---
title: "Quarto Assignment MT5000"
subtitle: "Shadows and Light: Mapping the Landscape of Child Deprivation Worldwide"

author: "Harshita Bhandari"
date: "2024-04-25"
format:
  html:
    theme: sandstone
    toc: true
    
execute: 
  echo: false
  warning: false
  message: false
---
```{r setup, include=FALSE}
#libraries
library(tidyverse)
library(ggplot2)
library(maps)
library(dplyr)
library(plotly)
library("png")
#data read
cleaned_unicef1 <- read.csv("cleaned_unicef1.csv")
cleaned_metadata <- read.csv("cleaned_metadata.csv")
total_data <- read.csv("total_data.csv")

#joining 
data_join <- full_join(cleaned_metadata, cleaned_unicef1)
data_join <- full_join(cleaned_metadata, cleaned_unicef1, by = join_by(country, year))
data_join <- full_join(cleaned_metadata, cleaned_unicef1, by = c("country" , "year" ))
full_join(cleaned_metadata, cleaned_unicef1, by = join_by(country, year))
full_join(cleaned_metadata, cleaned_unicef1, by = c("country" , "year" ))
```

## Insights from the World Map
The map displayed below uses a color gradient to denote the intensity of child deprivation. Darker shades indicate higher percentages of children experiencing at least two deprivations, signaling regions where child welfare needs urgent attention. This visual tool allows policymakers, researchers, and the public to identify focus areas for resource allocation and intervention efforts.
```{r}
#map
library(tidyverse)
library(ggplot2)
library(maps)
library(dplyr)
library(plotly)
map_world <- map_data("world")

map_data_join <- full_join(total_data, map_world, by = c("country" = "region"))                       
ggplot(map_data_join) +
  aes(x = long, y = lat, group = group, fill = obs_value) +
  geom_polygon() +
  labs(title = "Global Distribution of Child Deprivation ",  
subtitle = "Percentage of children suffering at least two deprivations")
```
## Insights from the Bar Graph

The visualization below illustrates the top ten countries facing severe child deprivation issues. Each bar represents a country and is color-coded to distinguish between them easily. The height of the bar indicates the average percentage of children suffering from at least two types of deprivation, providing a clear comparison of the severity across these nations.

By observing the data, we can see significant variation in child welfare across different regions. Countries with higher bars indicate a greater average percentage of children facing multiple deprivations, suggesting a higher priority for intervention in these areas.
```{r}
# bar graph
average_deprivation <- cleaned_unicef1 %>%
  group_by(country) %>%
  summarize(AverageDeprivation = mean(obs_value, na.rm = TRUE))
top_countries <- average_deprivation %>%
  arrange(desc(AverageDeprivation)) %>%
  top_n(10, AverageDeprivation) %>%
  mutate(country = factor(country, levels = rev(unique(country))))  # Ensure correct ordering in the plot

ggplot(top_countries, aes(x = country, y = AverageDeprivation, fill = country)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(  # Adjust y-axis scale to fit the data range properly
    labels = scales::percent_format(accuracy = 1)
  ) +
  scale_fill_viridis_d(direction = -1, guide = "none") +  # Color scale
  labs(title = "A Closer Look at Inequality: Child Welfare by Country",
       subtitle = "Top 10 Countries by Average Percentage of Children Suffering from At Least Two Deprivations 2011-2018",
       x = "Country",
       y = "Average Percentage of Children in Deprivation") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12),  # Improve legibility of country names
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    legend.position = "none"  # Removing the legend as it's redundant
  )
```
## Scatter Plot Explanation

The scatter plot below visualizes the relationship between military expenditure as a percentage of GDP and GDP per capita for the selected countries. Each point on the plot represents a data point for one of the countries in a given year, allowing us to observe trends over time and across different levels of economic development.

## Analysis and Discussion

The regression line in the scatter plot suggests a trend where countries with higher military expenditure as a percentage of GDP tend to have lower GDP per capita. This relationship might indicate that excessive military spending could be at the expense of economic growth, especially in nations where development funds are limited. However, it's also crucial to consider that high defense spending can be a response to security concerns, which might itself be a reaction to or cause of economic challenges.

## Recommendations

Given the observed trend, countries should evaluate the efficiency and necessity of their military budgets, especially when economic development goals remain unmet. Strategic reallocation of funds could support health, education, and infrastructure, potentially boosting economic growth more effectively than equivalent spending on defense. International cooperation and regional stability initiatives can also reduce security threats, decreasing the need for high military expenditure.

Policymakers should also consider comprehensive economic reforms that balance security needs with economic development, leveraging detailed data analyses like this to guide decisions.
```{r}
#scatter plot

filtered_data <- data_join %>%
  filter(
    country %in% c( "India", "Bangladesh", "Pakistan", "Indonesia", "Philippines")) %>%
  group_by(country)
# Create scatter plot with regression line
scatter_plot <- ggplot(filtered_data, aes(x = Military.expenditure....of.GDP., y = GDP.per.capita..constant.2015.US..)) +
  geom_point() +  # Scatter plot
  geom_smooth(method = "lm", se = FALSE) +  # Add regression line without confidence interval
  labs(x = "Military Eexpenditure as % of GDP", y = "GPD Per Capita", title = "Scatter Plot with Regression Line") +  # Labels
  theme_minimal()  # Optional: Customize plot theme

# Convert to plotly for an interactive plot
plotly_scatter_plot <- ggplotly(scatter_plot, tooltip = "text")

# If you're in an R environment that supports it, you can directly display the plot
plotly_scatter_plot
```


