---
title: "Quarto Assignment MT5000"
subtitle: "Shadows and Light: Mapping the Landscape of Child Deprivation"

author: "Harshita Bhandari"
date: "2024-04-25"
format:
  html:
    theme: lumen
    toc: true
    
execute: 
  echo: false
  warning: false
  message: false
---
```{r setup, include=FALSE}
#libraries
library(tidyverse)
library(ggplot2)
library(maps)
library(dplyr)
library(plotly)
library("png")
#data read
cleaned_unicef1 <- read.csv("cleaned_unicef1.csv")
cleaned_metadata <- read.csv("cleaned_metadata.csv")
total_data <- read.csv("total_data.csv")

#joining 
data_join <- full_join(cleaned_metadata, cleaned_unicef1)
data_join <- full_join(cleaned_metadata, cleaned_unicef1, by = join_by(country, year))
data_join <- full_join(cleaned_metadata, cleaned_unicef1, by = c("country" , "year" ))
full_join(cleaned_metadata, cleaned_unicef1, by = join_by(country, year))
full_join(cleaned_metadata, cleaned_unicef1, by = c("country" , "year" ))
```

## 1. Insights from the World Map
The map displayed below uses a color gradient to denote the intensity of child deprivation. Darker shades indicate higher percentages of children experiencing at least two deprivations, signaling regions where child welfare needs urgent attention. This visual tool allows policymakers, researchers, and the public to identify focus areas for resource allocation and intervention efforts.
```{r}
#map
library(tidyverse)
library(ggplot2)
library(maps)
library(dplyr)
library(plotly)
map_world <- map_data("world")

map_data_join <- full_join(total_data, map_world, by = c("country" = "region"))                       
ggplot(map_data_join) +
  aes(x = long, y = lat, group = group, fill = obs_value) +
  geom_polygon() +
  labs(title = "Global Distribution of Child Deprivation ",  
subtitle = "Percentage of children suffering at least two deprivations")
```
## 2. Insights from the Bar Graph

The visualization below illustrates the top ten countries facing severe child deprivation issues. Each bar represents a country and is color-coded to distinguish between them easily. The height of the bar indicates the average percentage of children suffering from at least two types of deprivation, providing a clear comparison of the severity across these nations.

By observing the data, we can see significant variation in child welfare across different regions. Countries with higher bars indicate a greater average percentage of children facing multiple deprivations, suggesting a higher priority for intervention in these areas.
```{r}
# bar graph
average_deprivation <- cleaned_unicef1 %>%
  group_by(country) %>%
  summarize(AverageDeprivation = mean(obs_value, na.rm = TRUE))
top_countries <- average_deprivation %>%
  arrange(desc(AverageDeprivation)) %>%
  top_n(10, AverageDeprivation) %>%
  mutate(country = factor(country, levels = rev(unique(country))))  # Ensure correct ordering in the plot

ggplot(top_countries, aes(x = country, y = AverageDeprivation, fill = country)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(  # Adjust y-axis scale to fit the data range properly
    labels = scales::percent_format(accuracy = 1)
  ) +
  scale_fill_viridis_d(direction = -1, guide = "none") +  # Color scale
  labs(title = "A Closer Look at Inequality: Child Welfare by Country",
       subtitle = "Top 10 Countries by Average Percentage of Children Suffering from At Least Two Deprivations 2011-2018",
       x = "Country",
       y = "Average Percentage of Children in Deprivation") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12),  # Improve legibility of country names
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    legend.position = "none"  # Removing the legend as it's redundant
  )
```


## 3. Scatter Plot Explanation

The scatter plot below visualizes the relationship between military expenditure as a percentage of GDP and GDP per capita for the selected countries. Each point on the plot represents a data point for one of the countries in a given year, allowing us to observe trends over time and across different levels of economic development.

## Analysis and Discussion

The scatter plot with regression lines for each country, the visualization uncovers patterns and raises questions:

Ethiopia:
Ethiopia, a country with historical conflicts and recent economic expansion, showcases a pattern of military spending reflective of its need to maintain national security while propelling forward its developmental agenda.

Madagascar:
Madagascar's unique position as an island nation with limited regional conflicts allows it to potentially prioritize economic development over military expansion, a hypothesis that its position on the scatter plot may support.

Cambodia:
Cambodia, with its post-conflict rebuilding phase and growing economic presence in Southeast Asia, presents an intriguing case of balancing military modernization with significant international investment in development.

Chad:
Chad's involvement in regional security initiatives in the face of economic challenges presents a complex interplay between defense commitments and the pursuit of economic stability.

Niger:
Niger faces unique security challenges with its strategic location and the threat of extremism. Its military expenditure reflects the delicate balance between ensuring security and striving towards economic development.

Given the observed trend, countries should evaluate the efficiency and necessity of their military budgets, especially when economic development goals remain unmet. Strategic reallocation of funds could support health, education, and infrastructure, potentially boosting economic growth more effectively than equivalent spending on defense. International cooperation and regional stability initiatives can also reduce security threats, decreasing the need for high military expenditure.

Policymakers should also consider comprehensive economic reforms that balance security needs with economic development, leveraging detailed data analyses like this to guide decisions.


```{r}

filtered_data <- data_join %>%
  filter(country %in% c("Ethiopia", "Madagascar", "Cambodia", "Chad", "Niger"))

# Create scatter plot with regression line with different colors for each country
scatter_plot <- ggplot(filtered_data, aes(x = Military.expenditure....of.GDP., y = GDP.per.capita..constant.2015.US.., color = country)) +
  geom_point() +  # Scatter plot
  geom_smooth(method = "lm", se = FALSE, aes(group = country)) +  # Add regression line without confidence interval for each country
  labs(x = "Military Expenditure as % of GDP", y = "GPD Per Capita", title = "Balancing Act: Military Investment and Economic Prosperity in Economies") +  # Labels
  theme_minimal() +  # Optional: Customize plot theme
  theme(legend.position = "bottom")  # Adjust legend position

# Convert to plotly for an interactive plot
plotly_scatter_plot <- ggplotly(scatter_plot, tooltip = c("x", "y", "color"))

# If you're in an R environment that supports it, you can directly display the plot
plotly_scatter_plot
```

## 4. Visualizing Population Trends:

To further enrich our understanding, we now look at a time series visualization of the population totals for these countries. By observing population growth trends alongside child deprivation data, we can better comprehend the scale of the challenges faced and the progress made over time.

This interactive plot allows viewers to trace the population changes over time, providing context to the severity and persistence of child deprivation issues. As the population increases, the demand for essential services such as education, healthcare, and economic opportunities also grows. The challenge is not only to reduce the percentage of children facing deprivation but also to scale these efforts effectively against a backdrop of rapid population growth.

```{r}
cleaned_metadata <- read.csv("cleaned_metadata.csv")
    
    # Filter data for specific countries
    data_selected <- subset(cleaned_metadata, country %in% c("Ethiopia", "Madagascar", "Cambodia", "Chad", "Niger"))
    
    # Create the ggplot object for time series data
    ggplot_object <- ggplot(data_selected, aes(x = year, y = Population..total, color = country, group = country)) +
      geom_line() +
      labs(x = "Year", y = "Population", title = "Population Growth Trends in Key Developing Nations: A Comparative Analysis") +
      theme_minimal()
    
    # Convert ggplot to an interactive Plotly graph
    plotly_object <- ggplotly(ggplot_object)
    
    # Display the interactive plot
    plotly_object
```

## 5. Conclusion
The insights garnered from these diverse analyses provided a strong empirical foundation for targeted policy recommendations. It became evident that a reallocation of resources—from predominantly military to more social services—could enhance child welfare significantly. Furthermore, these policies need to consider demographic trends to ensure that interventions are not only effective today but also sustainable in the long run.

This multi-faceted analysis demonstrated the importance of a balanced approach in policy-making, where economic growth, security, and social welfare are integrated into a coherent strategy. For the countries analyzed, effective policy-making will require an alignment of economic reforms with proactive social policies that collectively address the immediate and future needs of their populations. Moving forward, leveraging data-driven insights will be crucial in crafting policies that are both responsive and responsible, ensuring that children in these regions have the opportunity to live healthier, more prosperous lives in a stable environment. This holistic approach is not just beneficial but necessary for fostering sustainable development and ensuring that no child is left behind as nations progress.

## 6. Recommendations

The comprehensive analysis presented offers a multi-dimensional view into the dynamics of child deprivation, economic decision-making, and demographic trends in selected developing countries. Through a series of visualizations and data interpretations, we've underscored the significant disparities in child welfare and the impacts of fiscal choices on social services. This deep dive into the interplay between military expenditures, economic growth, and social welfare culminates in actionable policy recommendations aimed at realigning resources towards enhancing the well-being of the most vulnerable populations. It is clear that a balanced, integrated approach in policy-making that harmonizes economic strategies with social development objectives is imperative. By embracing this methodology, these nations can ensure sustainable development and create an environment where no child is left behind, paving the way for a future where all children can thrive.












